### 2007-01-14 Documentation/part1.tex
JK - Minor corrections in chapters 1 - 3

correlation with every parameter except from $\deltacp$, \ie,

@@

This corresponds to projection onto the $\stheta$-$\deltacp$-plane:

@@

Let us now locate the exact position\footnote{For a exact definition of 

@@

Before one can use \GLOBES , one has to initialize the \GLOBES\
library \index{norm}{libglobes@{\tt libglobes}}: \index{norm}{Initialization!\GLOBES\ library}
\begin{function}
\index{norm}{Initialization!libglobes@{\tt libglobes}}
\GLBNS{Init} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
to be called in the beginning of each \GLOBES\ program. It takes the
name {\tt name} of the program as a string to initialize the error handling
functions. In many cases, it is sufficient to use the first
argument from the command line as the program name (such as in example on page~\pageref{ex:c}).
\end{function}

\index{norm}{Experiment!number of}
In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
of existing experiment definition files can, of course, be unlimited. This means that their $\Delta \chi^2$-values are added {\em after} the minimization over the systematics parameters, and {\em before} any minimization over the oscillation parameters. Note that each experiment
assumes a specific matter density profile, which means
that it makes sense to simulate different operation modes within one
experiment definition, and physically different baselines in different
definitions. For details of the rate computation and
simulation techniques, we refer at this place to \Part~\ref{part:2}. Though
 the simplest case of simulating one experiment may be most often used, 
 using more than one experiments are useful in many cases. For example, combinations of experiments can be tested for
complementarity and competitiveness by equal means within one program.
In general, many \GLOBES\ functions take the experiment number as
a parameter, which runs from $0$ to \GLB{\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
the global variable {\tt glb\_num\_of\_exps} must not be modified by the
user.} In addition, using the parameter value \GLBC{GLB\_ALL} as
experiment number (for example, in the {\tt glbChi...} functions) initiates a combined analysis of all loaded experiments.
\index{norm}{GLB ALL@{\tt GLB\_ALL}}
\index{constants}{{\tt GLB\_ALL}}

@@

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$
as well as some other information, the parameter type \GLB{\_params} is used. In general,
 this type is often
transferred to and from \GLOBES\ functions. Therefore, the memory 
for these vectors has to be reserved (allocated) before they can be used, 
and it has
to be returned (freed) afterwards. \GLOBES\ functions usually use the pointers
of the type {\tt glb\_params} for the input or output to the functions. As
an input parameter, the pointer has to be created and point towards a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, the pointer has to be created, too, and point towards
a structure which the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the pointer structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{norm}{Matter density!scaling factor}
\index{norm}{Matter density!profile} 
which is stored in the density information of {\tt glb\_params}.
% CHANGED IN GLOBES 3.0: 
% Each of
% these scaling factors has $1.0$ as pre-defined value. Since it
% is in most cases not necessary to change this value, the user does
% not need to take care of it. 
For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, freed and allocated again.

Another piece of information will be returned from the minimizers 
(\cf, \Chapt~\ref{chapt:correlations}) and transferred
into the {\tt glb\_params} structure is the number
of iterations used for the minimization, which is proportional to the
running time of the minimizer. In general, the user does not need to
access the elements in {\tt glb\_params} directly. A number of functions
is provided to handle these parameter structures:
\begin{function}
\GLBNS{AllocParams}
{\tt glb\_params glbAllocParams()} allocates the memory space
needed for a parameter vector and returns a pointer to it. All parameter values are
initially set to {\tt nan} (not a number).
\end{function}
\begin{function}
\GLBNS{AllocParams}
{\tt void glbFreeParams(glb\_params stale)} frees the memory
needed for a parameter vector {\tt stale} and sets the pointer to {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{DefineParams}
{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dm21, double dm31)} assigns
the complete set of oscillation parameters to the vector {\tt in}, which has
to be allocated before. The return value is the pointer to {\tt in} if 
the assignment was successful, and {\tt NULL} otherwise.
\end{function}
\begin{function}
 \GLBNS{CopyParams}
 {\tt glb\_params glbCopyParams(const glb\_params source, glb\_params dest)}
 copies the vector {\tt source} to the vector {\tt destination}. The return
 value is {\tt NULL} if the assignment was not successful.
\end{function}
\begin{function}
\GLBNS{PrintParams}
{\tt void glbPrintParams(FILE *stream, const glb\_params in)} prints the
parameters in {\tt in} to the file {\tt stream}. The oscillation
parameters, all density values, and the number of iterations are
printed as pretty output. Use {\tt stdout} for {\tt stream} if you want
to print to the screen.
\end{function}
In addition to these basic functions, there are functions to access
the individual parameters within the parameter vectors:
\begin{function}
\GLBNS{SetOscParams}
{\tt glb\_params glbSetOscParams(glb\_params in, double osc, int which)}
sets the oscillation parameter {\tt which} in the structure {\tt in}
to the value {\tt osc}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{GetOscParams}
{\tt double glbGetOscParams(glb\_params in, int which)}
returns the value of the oscillation parameter {\tt which} in the 
structure {\tt in}.
\end{function}
In both of these functions, the parameter {\tt which} runs from
$0$ to $5$ (or the number of oscillation parameters-1), where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
\GLBC{GLB\_DM\_21}, or \GLBC{GLB\_DM\_31} can be used.

@@

Compared to existing experiments, which use real data, future experiments
use simulated data. Thus, the {\em true parameter values} and their results in 
form of the reference event rate vectors are simulated. After setting the 
true parameter values, the {\em fit parameter values} can be varied in order to 
obtain information on the measurement performance for the given set of true 
parameter values. Therefore, it is often useful to show the results of a 
future measurement as function of the true parameter values for which the 
reference rate vectors are computed -- at least within the currently allowed 
ranges. The true parameter values for the vacuum neutrino oscillation 
parameters have to be set by the function \GLB{SetOscillationParameters} 
and the reference rate vector, \ie\ the data, has to be computed by a call 
to \GLB{SetRates}.
This has to be done {\em before} any evaluation 
function is used and  {\em after} the experiments have been initialized and 
also the experiment 
parameters have been adjusted which could change the rates (such as baseline
 or target mass). This means that
after any change of an experiment parameter, \GLB{SetRates} has to be 
called. Matter effects are automatically included as specified in the 
experiment definition. We have the following functions to assign and
read out the vacuum oscillation parameters:
\begin{function}
\GLBNS{SetOscillationParameters}
{\tt int glbSetOscillationParameters(const glb\_params in)} sets the
vacuum oscillation parameters to the ones in the vector {\tt in}.
\end{function}
\begin{function}
\GLBNS{GetOscillationParameters}
{\tt int glbGetOscillationParameters(glb\_params out)} returns the
vacuum oscillation parameters in the vector {\tt out}. The result of
the function is $0$ if the call was successful.
\end{function}
The reference rate vector is then computed with:
\begin{function}
\GLBNS{SetRates}
{\tt void glbSetRates()} computes the reference rate vector for the neutrino oscillation parameters set by {\tt glbSetOscillationParameters}. 
\end{function}
A complete example for a minimal \GLOBES\ program can be found on
Page~\pageref{ex:c}.

@@

\index{norm}{Systematics} \index{norm}{Pull method}
The treatment of systematics in \GLOBES\ is performed by the so-call
{\em pull method} with the help of auxiliary systematics parameters.
They are taken completely uncorrelated among different rules, and treated with simple Gau\ss ian statistics. In general, a rule is a set 
of signal and background event rates coming from different oscillation
channels, where the event rates of all rule contributions are added.
For more details of the rule concept, see \Part~\ref{part:2} of this manual,
and for the treatment of systematics, see \Sec~\ref{sec:rules}.

 One example for a systematics parameter the signal normalization error, \ie, an error to the overall normalization of the signal. For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter\index{norm}{Auxiliary parameter} of this channel, \ie , 
\be
 s_i = s_i(n_s) = s_i^0 \cdot (1 + n_s),
\ee
where $n_s$ is the signal normalization parameter. The total number of events in the $i$th bin $x_i$ also includes the background event rates $b_i$, \ie, $x_i = s_i + b_i$, which may have their own systematics parameters.
In order to implement an overall signal normalization error $\sigma_{n_s}$,  the $\chi^2$, which includes all event rates $x_i$ of all bins, is minimized over the auxiliary parameter $n_s$:
\be
 \hat{\chi^2} = \underset{n_s}{\mathrm{min}} \left(  \chi^2(n_s, \hdots) + \frac{(n_s)^2}{\sigma_{n_s}^2} \right).
\ee 
This minimization is done independently for all auxiliary parameters of the rule. The total $\chi^2$ for the considered experiment is finally obtained by repeating this procedure for all rules and adding their $\chi^2$-values. In general, the situation is more complicated because of the usage of many systematical errors. More details about systematics parameters and the definition of signal, background, and oscillation channels can be found in \Sec~\ref{sec:rules}, too.

@@

the AEDL file defines user-defined systematics for a specific rule and gives it a 
name/identifier,  as will as it defines the systematical errors. The source code has then to match this AEDL

@@

Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates. The array indices are running from $0$ to {\tt dim}-1.

@@

Therefore the function should be as efficient as possible. In many cases, it is also recommendable to use different functions for different rules, since the fewer auxiliary parameters in the minimization, the faster the routine will be. In addition, note that complicated systematics with many parameters may introduce complicated (unknown) topologies for the minimization, which means that the minimizers may

@@

    \hspace*{0.5cm} for(i=0;i<dim;i++) chi2 += (x[i]*x[i]/errors[i]/errors[i]);

@@

During running time, it is also possible to change the user-defined systematics of an experiment or

@@

user-defined systematics can be used for each experiment or each rule. The parameter 
{\tt on\_off} determines if the systematics function should be used for systematics on
{\tt GLB\_ON} or systematics off {\tt GLB\_OFF}.
The array {\tt errors} sets the systematical errors in the order they
are defined in the AEDL file (running from $0$ to the number of systematics parameters-1). 
Returns zero, if successful.

@@

Note that the new user-defined systematics has to be registered with {\tt glbDefineChiFunction} first. 
One can also return the systematics function name by 

@@

The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
for the $\chi^2$ calculation. 
A very useful function to include energy calibration errors is
\begin{function}
\GLBNS{ShiftEnergyScale}
{\tt glbShiftEnergyScale(double b, double *rates\_in, double *rates\_out,int bins, double emin, double emax)}
shifts the energy scale in {\tt rates\_in} by the amount {\tt b} and stores the result in 
{\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
as obtained with \GLB{GetEminEmax}. Requires constant energy bin widths!
\end{function}
For details on energy calibration errors, see \Sec~\ref{sec:rules}.
%
In addition, in order to implement user-defined systematics, one usually needs low-level access to the event rate vectors using \GLB{GetRuleRatePtr}, \GLB{GetSignalFitRatePtr}, \GLB{GetBGFitRatePtr} and possibly other functions described in \Sec~\ref{sec:event_rates}. Additionally, the
functions \GLB{GetEminEmax}, \GLB{GetEnergyWindow}, \GLB{GetEnergyWindowBins}, and \GLB{GetNumberOfBins} from \Sec~\ref{sec:aedl_names} are useful.  We recommend that you familiarize yourself with these functions at this point.

>>>>

correlation with every parameter except for $\deltacp$, \ie,

@@

This corresponds to a projection onto the $\stheta$-$\deltacp$-plane:

@@

Let us now locate the exact position\footnote{For an exact definition of 

@@

Before one can use \GLOBES , one has to initialize the \GLOBES\
library \index{norm}{libglobes@{\tt libglobes}}: \index{norm}{Initialization!\GLOBES\ library}
\begin{function}
\index{norm}{Initialization!libglobes@{\tt libglobes}}
\GLBNS{Init} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
to be called in the beginning of each \GLOBES\ program. It takes the
name {\tt name} of the program as a string to initialize the error handling
functions. In many cases, it is sufficient to use the first
argument from the command line as the program name (such as in the example on page~\pageref{ex:c}).
\end{function}

\index{norm}{Experiment!number of}
In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
of existing experiment definition files can, of course, be unlimited. This means that their $\Delta \chi^2$-values are added {\em after} the minimization over the systematics parameters, and {\em before} any minimization over the oscillation parameters. Note that each experiment
assumes a specific matter density profile, which means
that it makes sense to simulate different operation modes within one
experiment definition, and physically different baselines, in different
definitions. For details of the rate computation and
simulation techniques, we refer at this place to \Part~\ref{part:2}. Though
 the simplest case of simulating one experiment may be most often used, 
 using several experiments is useful in many cases. For example, combinations of experiments can be tested for
complementarity and competitiveness by equal means within one program.
In general, many \GLOBES\ functions take the experiment number as
a parameter, which runs from $0$ to \GLB{\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
the global variable {\tt glb\_num\_of\_exps} must not be modified by the
user.} In addition, using the parameter value \GLBC{GLB\_ALL} as
experiment number (for example, in the {\tt glbChi...} functions) initiates a combined analysis of all loaded experiments.
\index{norm}{GLB ALL@{\tt GLB\_ALL}}
\index{constants}{{\tt GLB\_ALL}}

@@

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$
as well as some other information, the parameter type \GLB{\_params} is used. In general,
 this type is often
transferred to and from \GLOBES\ functions. Therefore, the memory 
for these vectors has to be reserved (allocated) before they can be used, 
and it has to be returned (freed) afterwards.
\GLOBES\ functions usually use pointers
of the type {\tt glb\_params} for the input or output to the functions.
As an input parameter, the pointer has to be initialized with the address of a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, it has to be initialized with the address of
a structure which the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the \GLB{\_params} structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{norm}{Matter density!scaling factor}
\index{norm}{Matter density!profile} 
which is stored in the density information of {\tt glb\_params}.
% CHANGED IN GLOBES 3.0: 
% Each of
% these scaling factors has $1.0$ as pre-defined value. Since it
% is in most cases not necessary to change this value, the user does
% not need to take care of it. 
For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, freed and allocated again.

Another piece of information that will be returned from the minimizers 
(\cf, \Chapt~\ref{chapt:correlations}) and transferred
into the {\tt glb\_params} structure is the number
of iterations used for the minimization, which is proportional to the
running time of the minimizer. In general, the user does not need to
access the elements in {\tt glb\_params} directly. A number of functions
is provided to handle these parameter structures:
\begin{function}
\GLBNS{AllocParams}
{\tt glb\_params glbAllocParams()} allocates the memory space
needed for a parameter vector and returns a pointer to it. All parameter values are
initially set to {\tt nan} (not a number).
\end{function}
\begin{function}
\GLBNS{AllocParams}
{\tt void glbFreeParams(glb\_params stale)} frees the memory
needed for a parameter vector {\tt stale} and sets the pointer to {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{DefineParams}
{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dm21, double dm31)} assigns
the complete set of oscillation parameters to the vector {\tt in}, which has
to be allocated before. The return value is the pointer to {\tt in} if 
the assignment was successful, and {\tt NULL} otherwise.
\end{function}
\begin{function}
 \GLBNS{CopyParams}
 {\tt glb\_params glbCopyParams(const glb\_params source, glb\_params dest)}
 copies the vector {\tt source} to the vector {\tt destination}. The return
 value is {\tt NULL} if the assignment was not successful.
\end{function}
\begin{function}
\GLBNS{PrintParams}
{\tt void glbPrintParams(FILE *stream, const glb\_params in)} prints the
parameters in {\tt in} to the file {\tt stream}. The oscillation
parameters, all density values, and the number of iterations are
printed as pretty output. Use {\tt stdout} for {\tt stream} if you want
to print to the screen.
\end{function}
In addition to these basic functions, there are functions to access
the individual parameters within the parameter vectors:
\begin{function}
\GLBNS{SetOscParams}
{\tt glb\_params glbSetOscParams(glb\_params in, double osc, int which)}
sets the oscillation parameter {\tt which} in the structure {\tt in}
to the value {\tt osc}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{GetOscParams}
{\tt double glbGetOscParams(glb\_params in, int which)}
returns the value of the oscillation parameter {\tt which} in the 
structure {\tt in}.
\end{function}
In both of these functions, the parameter {\tt which} runs from
$0$ to $5$ (or the number of oscillation parameters-1), where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
\GLBC{GLB\_DM\_21}, or \GLBC{GLB\_DM\_31} can be used.

@@

Compared to existing experiments, which use real data, future experiments
use simulated data. Thus, the {\em true parameter values} and their results in 
form of the reference event rate vectors are simulated. After setting the 
true parameter values, the {\em fit parameter values} can be varied in order to 
obtain information on the measurement performance for the given set of true 
parameter values. Therefore, it is often useful to show the results of a 
future measurement as function of the true parameter values for which the 
reference rate vectors are computed -- at least within the currently allowed 
ranges. The true parameter values for the vacuum neutrino oscillation 
parameters have to be set by the function \GLB{SetOscillationParameters}, 
and the reference rate vector, \ie\ the data, has to be computed by a call 
to \GLB{SetRates}.
This has to be done {\em before} any evaluation 
function is used and  {\em after} the experiments have been initialized and 
also the experiment 
parameters have been adjusted which could change the rates (such as baseline
 or target mass). This means that
after any change of an experiment parameter, \GLB{SetRates} has to be 
called. Matter effects are automatically included as specified in the 
experiment definition. We have the following functions to assign and
read out the vacuum oscillation parameters:
\begin{function}
\GLBNS{SetOscillationParameters}
{\tt int glbSetOscillationParameters(const glb\_params in)} sets the
vacuum oscillation parameters to the ones in the vector {\tt in}.
\end{function}
\begin{function}
\GLBNS{GetOscillationParameters}
{\tt int glbGetOscillationParameters(glb\_params out)} returns the
vacuum oscillation parameters in the vector {\tt out}. The result of
the function is $0$ if the call was successful.
\end{function}
The reference rate vector is then computed with:
\begin{function}
\GLBNS{SetRates}
{\tt void glbSetRates()} computes the reference rate vector for the neutrino oscillation parameters set by {\tt glbSetOscillationParameters}. 
\end{function}
A complete example for a minimal \GLOBES\ program can be found on
Page~\pageref{ex:c}.

@@

\index{norm}{Systematics} \index{norm}{Pull method}
The treatment of systematics in \GLOBES\ is performed by the so-called
{\em pull method} with the help of auxiliary systematics parameters.
They are taken to be completely uncorrelated among different rules,
and treated with simple Gau\ss ian statistics.
In general, a rule is a prescription for summing up experimentally indistinguishable
signal and background events from different oscillation channels.
For more details on the rule concept, see \Part~\ref{part:2} of this manual,
and for the treatment of systematics, see \Sec~\ref{sec:rules}.

 One example for a systematics parameter is the signal normalization error, \ie, an error on the overall normalization of the signal. For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter\index{norm}{Auxiliary parameter} of this channel, \ie , 
\be
 s_i = s_i(n_s) = s_i^0 \cdot (1 + n_s),
\ee
where $n_s$ is the signal normalization parameter. The total number of events in the $i$th bin $x_i$ also includes the background event rates $b_i$, \ie, $x_i = s_i + b_i$, which may have their own systematics parameters.
In order to implement an overall signal normalization error $\sigma_{n_s}$,  the $\chi^2$, which includes all event rates $x_i$ of all bins, is minimized over the auxiliary parameter $n_s$:
\be
 \hat{\chi^2} = \underset{n_s}{\mathrm{min}} \left(  \chi^2(n_s, \hdots) + \frac{(n_s)^2}{\sigma_{n_s}^2} \right).
\ee 
This minimization is done independently for all auxiliary parameters of the rule. The total $\chi^2$ for the considered experiment is finally obtained by repeating this procedure for all rules and adding their $\chi^2$-values. In general, the situation is more complicated because of the usage of many systematical errors. More details about systematics parameters and the definition of signal, background, and oscillation channels can be found in \Sec~\ref{sec:rules}, too.

@@

the AEDL file defines user-defined systematics for a specific rule, gives it a 
name/identifier, and defines the systematical errors. The source code has then to match this AEDL

@@

Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates. The array indices run from $0$ to {\tt dim}-1.

@@

Therefore the function should be as efficient as possible.
%JK - THIS IS INCORRECT: ONE MINIMIZATION OVER, SAY, 8 PARAMETERS IS FASTER THAN
%2 MINIMIZATIONS OVER 4 PARAMETERS EACH-
%In many cases, it is also recommendable to use different functions for different
%rules, since the fewer auxiliary parameters in the minimization, the faster the routine will be.
In addition, note that complicated systematics with many parameters may introduce complicated (unknown) topologies for the minimization, which means that the minimizers may

@@

    \hspace*{0.5cm} for(i=0;i<dim;i++) chi2 += (x[i]*x[i])/(errors[i]*errors[i]);

@@

During running time, it is also possible to change the systematics of an experiment or

@@

given systematics function should be used for each experiment or each rule. The parameter 
{\tt on\_off} determines if the systematics function should be used when systematical errors
are switched on ({\tt GLB\_ON}), or when they are switched off ({\tt GLB\_OFF}).
The array {\tt errors} sets the systematical errors in the order in which they
are expected by the systematics function (indices run from $0$ to
the number of systematics parameters-1). The function returns zero, if successful.

@@

Note that new user-defined systematics functions have to be registered with {\tt glbDefineChiFunction} first. 
One can also request the systematics function name by 

@@

The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
for the $\chi^2$ calculation. 
A very useful function to include energy calibration errors is
\begin{function}
\GLBNS{ShiftEnergyScale}
{\tt glbShiftEnergyScale(double b, double *rates\_in, double *rates\_out,int bins, double emin, double emax)}
shifts the energy scale in {\tt rates\_in} by the relative amount {\tt b} and stores the result in 
{\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
as obtained with \GLB{GetEminEmax}. Requires constant energy bin widths!
\end{function}
For details on energy calibration errors, see \Sec~\ref{sec:rules}.
%
In addition, in order to implement user-defined systematics, one usually needs low-level access to the event rate vectors using \GLB{GetRuleRatePtr}, \GLB{GetSignalFitRatePtr}, \GLB{GetBGFitRatePtr} and possibly other functions described in \Sec~\ref{sec:event_rates}. Additionally, the
functions \GLB{GetEminEmax}, \GLB{GetEnergyWindow}, \GLB{GetEnergyWindowBins}, and \GLB{GetNumberOfBins} from \Sec~\ref{sec:aedl_names} are useful.  We recommend that you familiarize yourself with these functions at this point.

Let us now discuss a simple example.  Typical applications for user-defined systematics are reactor neutrino experiments, where systematics play a crucial role. For a setup with near and far detectors,
the formula for $\chi^2$ is
\begin{align}
  \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F}2
     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} +
       (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) \nonumber\\
         & \ \ + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
\label{equ:chi2reactor}
\end{align}
where $O_{N,i}$ and $O_{F,i}$ are the event rates for the $i$-th bin in the near and
far detector, calculated for the assumed ``true'' values of the oscillation parameters.
$T_{d,i}$ are the event rates for the parameter values that are currently being tested.
$a_R$, $a_N$ and $a_F$ parametrize the small uncertainties in the reactor flux and the fiducial
mass of the two detectors. In this example, their central values are assumed to be zero,
while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
%
The first line of \equ{chi2reactor} is the standard $\chi^2$ expression for
the Poisson distribution, while the terms in the second line are penalties
for deviations of the systematics parameters from their central values. 
%
We use two AEDL files for this experiment, one for the far detector and one for the
near detector.
The corresponding rule definition ($\nu_e$ disappearance) in the far detector AEDL-file is
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiDCNorm" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ 0.02,    0.006,        0.006,       0.005,     0.005   \} \\
\hspace*{0.5cm}        \mbox{ // \{ Flux, Fid. mass FD, Fid. mass ND, Energy FD, Energy ND \} } \\
> 
}
\end{quote}
In this case, the systematics is called ``{\tt chiDCNorm}'' for systematics on, whereas systematics
off is computed with spectral information, but without systematical errors.
For the near detector AEDL-file, we have instead
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiZero" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ \} \\
> 
}
\end{quote}
In this case, the systematics {\tt chiZero} is used for systematics on, which means that there will
be no active $\chi^2$ calculation in this rule. With this definition, the user-defined systematics
will only be called once for the far detector. However, the rates from the near detector will be 
passively provided for the common $\chi^2$ function.
You can find the corresponding files {\tt dchooz-near.glb} and {\tt dchooz-far.glb} you will need for {\tt example5.c}, in the example directory. 
%
We show the implementation of the $\chi^2$ function in the example on page~\pageref{ex:syst},
where we in addition include an uncorrelated energy calibration error. See \figu{reactor} for the result
of this example.
%
\begin{figure}[t]
\begin{center}
\includegraphics[width=10cm]{reactor}
\end{center}
\mycaption{\label{fig:reactor} The result from {\tt example5.c} (figure similar to the one
in \Ref~\cite{Huber:2006vr}). The thick curve corresponds to the systematics on page~\pageref{ex:syst}.}
\end{figure}

Further functions useful in the context of user-defined systematics are:
\begin{function}
\GLBNS{SetSysErrorsList} 
{\tt int glbSetSysErrorsList(int exp, int rule, int on\_off, const double *sys\_list)}
changes the systematical errors defined in the AEDL file in experiment {\tt exp} and 
rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
\end{function}
\begin{function}
\GLBNS{GetSysErrorsListPtr} 
{\tt double* glbGetSysErrorsListPtr(int exp, int rule, int on\_off)}
returns a pointer to the list of systematical errors as defined in the AEDL file in experiment {\tt exp} and 
rule {\tt rule}. Choose
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
\end{function}
In cases where the systematics minimization does not converge fast enough or ends
up in a local minimum, one can change the starting values of the minimizer, \ie, the
starting point from which the local minimizer rolls into the local minimum.
\begin{function}
\GLBNS{SetSysStartingValuesList} 
{\tt int glbSetSysStartingValuesList(int exp, int rule, int on\_off, const double *sys\_list)}
sets the starting values of the local systematics minimizer in experiment {\tt exp} and 
rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
Usually, zero is used for all starting values. 
\end{function}
\begin{function}
\GLBNS{GetSysStartingValuesListPtr} 
{\tt double *glbGetSysStartingValuesListPtr(int exp, int rule, int on\_off, const double *sys\_list)}
returns the starting values of the local systematics minimizer in experiment {\tt exp} and 
rule {\tt rule}. This list can be obtained for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}.
\end{function}
A useful trick is often to use the minimum from the last run as starting values for the next one
if the input parameters are only slighltly changed. One can obtain the minimum from the last run 
from the last call of the $\chi^2$ function (\cf, {\tt example5.c}).
